The complex network of trophic interactions in a subAntarctic Marine Protected

### 2023-01-31 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Add Discussion

In the subAntarctic region, the level of protection is mainly associated to oceanic islands, such as South Georgia and the South Sandwich, Bouvet, Prince Edward, and Macquarie [@IUCN2023]. Interestingly, the case of the MPA Namuncurá - Banco Burdwood I (MPA N-BB), which forms the focus of this work, is unique since it is associated to a submarine plateau [@Reta2014]. Added to this, such MPA is part of a network of protected areas in the subAntarctic (i.e joining MPAs Namuncurá - Banco Burdwood II and Yaganes), that contribute to protect and strengthen the functioning of this southern region to maintain global ocean health.

Many of these MPAs centre on the presence of particularly vulnerable, keystone, or charismatic species; the presence of large numbers (or proportions) of endemic species; and the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created because it is considered a biodiversity hotspot, mostly of its benthic community [@Schejter2016; @Falabella2017]. This community is featured by a high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge, or nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to the oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the area [@APN2022].

Recently, the identification of the most influential species has become a crucial task for conservation purposes in local and regional contexts to maintain ecosystem integrity and the supply of ecosystem services for the southwestern South Atlantic Ocean [@Franco2020; @Bergagna2020; @Riccialdelli2020].
In this regard, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g. fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other subAntarctic areas, like the Atlantic coast of Tierra del Fuego, where the Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are characterised by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @APN2022], there is a need for a better understanding of species interactions' complexity and structure.
This aspect can be tackled by analysing one of the most-frequent relationships occurring between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, which represents the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and respond to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Here it is worth mentioning that anthropogenic threats have been recently reported for the MPA N-BB region [@Cossi2021; @Fioramonti2022].

@@

The MPA Namuncurá - Banco Burdwood I is a shallow submarine plateau called Burdwood Bank (BB) located 150 km east of Isla de los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1). It comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. The BB is surrounded by steep flanks of more than 3000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. Intense upwelling and mixing occur over it, entraining deep nutrient rich waters into the photic layer [@Piola2009; @Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019]. Physical features in the MPAN-BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC overall [@Guerrero1999; @Acha2004; @Piola2009].

```{r figure1, echo=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Map of the Marine Protected Area Namuncurá - Banco Burdwood. Taken from https://www.argentina.gob.ar/parquesnacionales/areasmarinas/namuncura-burdwood."}
# Map of MPAN-BB provided by Lu R. (?)
knitr::include_graphics("MPABurdwood_map.jpeg")

@@

In order to build the network of predator-prey interactions we reviewed more than 150 references considering published articles, databases and doctoral theses. Furthermore, we took into account personal communications from experts belonging to the working group of the study area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the expertise of the authors contributing to the present study was a key factor in enhancing the quality of the network, and inherently improved the network representation of the MPA Namuncurá-Banco Burdwood ecosystem. A list of the references used to build the network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the MPA Namuncurá-Banco Burdwood, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases we followed this when data on specific biological species were not available, but for some cases we collapsed species when taxa shared the same set of predators and prey (trophic similarity criteria). Details about this can be found in Supplementary Material (Table S2).

@@

table1 <- add_row(table1, Name="Topological role", Definition="Species role according to interactions within and across modules (subgroups of species).", Implications="Four roles are defined: module hub, module specialist, module connector and network connector. Species with the same role are expected to have similar topological properties.", Reference="Guimera & Nunes Amaral 2005")

@@

The food web for the MPA Namuncurá - Banco Burdwood is one of the most highly-resolved networks of trophic interactions not only for high-latitude marine ecosystems but also for marine protected areas. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure of the ecosystem, since the aggregation of species might mask food web properties and/or generate artifacts [@Martinez1993; @Gauzens2013].

@@

The structure of species interactions influences the opportunities for multiple species to persist in the long term and, therefore, their coexistence [@Godoy2018].
The food web for the MPA N-BB ecosystem, as we showed in this study, presents an asymmetric degree (interaction) distribution. This is a pattern that we have identified at different levels of analysis: food web, predator, prey, and functional group. Such asymmetry is a well-known feature in empirical complex food webs in particular [@Dunne2002a; @Montoya2003; @Stouffer2005], and has received great attention in complex networks in general [@Albert2002; @Newman2003].
The degree distribution pattern affects the resilience of complex food webs against random failures and intentional attacks: more-skewed distributions, like the one described in this study, are more vulnerable to the removal of the most-connected species or hubs with the potential of producing a catastrophic fragmentation of the network [@Albert2000; @Dunne2002].

Considering the relevance of DD in network robustness we have considered this parameter in our analysis of food webs vulnerability to species loss. In particular we are interested in the intentional removal of certain species in the food web more than in the random extinction of species. As Dunne et al. (2002b) have correctly pointed out the anthropogenically caused species and population extinctions are far from being random as human have historically tended to impact higher trophic levels in the ecosystems.

@@

>>>>

In the subAntarctic region, the level of protection is mainly associated to oceanic islands, such as South Georgia and the South Sandwich, Bouvet, Prince Edward, and Macquarie [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB), which forms the focus of this work, is unique since they are associated to a submarine plateau and the southern adjacent deep slope region, respectively [@Falabella2017]. Added to this, such MPAs are part of a network of protected areas in the subAntarctic (jointly with MPA Yaganes), that contributes to protect and strengthen the functioning of this southern region to maintain global ocean health.

Many of these MPAs centre on the presence of particularly vulnerable, keystone, or charismatic species; the presence of large numbers (or proportions) of endemic species; and the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a sensitive and biodiverse benthic habitat that was barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by a high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge, or nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to the oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the area [@APN2022].

Recently, the identification of the most influential species has become a crucial task for conservation purposes in local and regional contexts to maintain ecosystem integrity and the supply of ecosystem services for the southwestern South Atlantic Ocean [@Franco2020; @Bergagna2020; @Riccialdelli2020].
In this regard, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g. fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other subAntarctic areas, like the Atlantic coast of Tierra del Fuego. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are characterised by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @APN2022], there is a need for a better understanding of species interactions' complexity and structure.
This aspect can be tackled by analysing one of the most-frequent relationships occurring between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, which represents the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and respond to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Here it is worth mentioning that anthropogenic threats have been recently reported for the MPA N-BB region [@Cossi2021; @Fioramonti2022].

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws 26875 and 27490, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, respectively [@APN2023]. They are located 150 km east of Isla de los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA Namuncurá - Banco Burdwood I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC overall [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. Such deep slope form the basis of the MPA Namuncurá - Banco Burdwood II that protects about 32,000 km^2^. Intense upwelling and mixing occur over it, entraining deep nutrient rich waters into the photic layer [@Piola2009; @Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

Given the evidence collected during several research cruises about the oceanographic and ecological processes connecting MPAs Namuncurá - Banco Burdwood I and II [references in @APN2022], a joint management plan was recently proposed [@APN2022]. Three categories were defined according to the authorized activities: 1) Marine National Reserve (MNR), an area with the objective of conserving marine biodiversity, landscape quality and large-scale ecological processes, allowing scientific, educational, recreational uses and the sustainable use of one or more of its natural resources; 2) Marine National Park, an area with the objective of conserving marine biodiversity, landscape quality and large-scale ecological processes (MNP); and 3) Restricted Marine National Reserve (RMNR), an area of maximum permanent or temporary protection with the objective to preserve marine biodiversity and ecological processes in the long term, minimizing anthropogenic impact, and allowing scientific research and monitoring, as well as control and surveillance activities. Figure 1 shows the geographic location of these categories within MPAs Namuncurá - Banco Burdwood I and II.

```{r figure1, echo=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Map of the Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve."}
# Map of MPAN-BB provided by Lu R. (?)
knitr::include_graphics("MPABurdwood_map.jpeg")

@@

In order to build the network of predator-prey interactions we reviewed more than 150 references considering published articles, databases and doctoral theses. Furthermore, we took into account personal communications from experts belonging to the working group of the study area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the expertise of the authors contributing to the present study was a key factor in enhancing the quality of the network, and inherently improved the network representation. A list of the references used to build the network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases we followed it when specific data on biological species were not available, but for some cases we collapsed species when taxa shared the same set of predators and prey (trophic similarity criterion, @Martinez1991). Details about this can be found in Supplementary Material (Table S2).

@@

table1 <- add_row(table1, Name="Topological role", Definition="Species role according to interactions within and across modules (subgroups of species).", Implications="Four roles are defined: module hub, module specialist, module connector and network connector. Network connector and module connector roles maintain the connectivity of the food web.", Reference="Guimera & Nunes Amaral 2005")

@@

The food web for the MPA Namuncurá - Banco Burdwood, analysed in this study, is one of the most highly-resolved networks of trophic interactions not only for a high-latitude marine ecosystem but also for a marine protected area worldwide. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure and functioning of the ecosystem, since the aggregation of species might mask food web properties and/or generate artifacts [@Martinez1993; @Gauzens2013].
Connectance, a property that resumes the complexity of a food web,

@@

The structure of species interactions influences the opportunities for multiple species to persist in the long term and, therefore, their coexistence [@Godoy2018].
The food web for the MPA N-BB ecosystem, as we showed in this study, presents an asymmetric degree (interaction) distribution. This is a pattern that we have identified at different levels of analysis: food web, predator, prey, and functional group. Such asymmetry is a well-known feature in empirical complex food webs in particular [@Dunne2002a; @Montoya2003; @Stouffer2005], and has received great attention in complex networks in general [@Albert2002; @Newman2003].
The degree distribution pattern affects the resilience of complex food webs against random failures and intentional attacks: more-skewed distributions, like the one described in this study, are more vulnerable to the removal of the most-connected species or hubs with the potential of producing a catastrophic fragmentation of the network [@Albert2000; @Dunne2002].
Important to attacks like anthropogenic perturbations: mercury, microplastics
As Dunne et al. (2002b) have correctly pointed out the anthropogenically caused species and population extinctions are far from being random as human have historically tended to impact higher trophic levels in the ecosystems.

The degree distribution allows identification of important species, such as potential keystone species [@Sole2001; @Dunne2002], generalists, and specialist species [@Kondoh2010].

Considering the relevance of DD in network robustness we have considered this parameter in our analysis of food webs vulnerability to species loss. In particular we are interested in the intentional removal of certain species in the food web more than in the random extinction of species.

@@

Incorporate habitat heterogeneity, since benthic community in the BB and deep flanks is different [@Schejter2021]. This might lead to distinct food web properties, hence differences in ecosystem functioning and response to anthropogenic and environmental perturbations.
@Cordone2020
